Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-10-23 prop_letter.tex
prop_letter.tex \section{Born rule and Gleason's Theorem} Updating...

\begin{definition}[QIVPM]\label{def:QIVPM}
  Assume a collection of intervals $\mathscr{I}$ including $\imposs$
  and $\necess$ with addition and scalar multiplication defined as
  follows:
  \begin{subequations}\label{eq:interval-operations}
  \begin{eqnarray}
   &  & [\ell_{1},r_{1}]+[\ell_{2},r_{2}]=[\ell_{1}+\ell_{2},r_{1}+r_{2}]\textrm{ and}\\
   &  & x[\ell,r]=\begin{cases}
  [x\ell,xr] & \textrm{for }x\ge0\,;\\{}
  [xr,x\ell] & \textrm{for }x\le0\,.
  \end{cases}
  \end{eqnarray}
  \end{subequations}
  Then we take a QIVPM~$\bar{\mu}$ to be an assignment of an interval to each
  event (projection operator~$P$) subject to the following constraints:
  \begin{subequations}\label{eq:QIVPM-constraints}
  \begin{eqnarray}
   &  & \bar{\mu}(\mathbb{0})=\imposs\,,\\
   &  & \bar{\mu}(\mathbb{1})=\necess\,,\label{eq:necess}\\
   &  & \bar{\mu}\left(\mathbb{1}-P\right)= \necess -\bar{\mu}\left(P\right)\,,
  \end{eqnarray}
  \end{subequations}
  and satisfying for each pair of \emph{commuting} projectors~$P_0$
  and~$P_1$ with $P_0P_1=P_1P_0$,
\begin{equation}
\bar{\mu}\left(P_{0}+P_{1}-P_{0}P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)\,.
\label{eq:QuantumInterval-valuedProbability-Inclusion}
\end{equation}
\end{definition}
\noindent The first three constraints,
Eqs.~(\ref{eq:QIVPM-constraints}), are the direct counterpart of the
corresponding ones for classical IVPMs.  With the understanding that
the union of classical sets $E_1\cup E_2$ is replaced by
$P_0+P_1-P_0P_1$ in the case of quantum projection
operators~\cite{Griffiths2003}, the last condition,
Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}), is a
direct counterpart of the convexity condition of
Eq.~(\ref{eq:classicalconvex}). Thus our definition of QIVPMs merges
classical IVPMs with quantum probability measures.

@@

Even if $\iota$ has to be a THOS rule to guarantee $\iota\circ\mu$
a QIVPM, the composited function $\iota\circ\mu$ may still be a QIVPM
when $\iota$ is not a THOS rule. In this case, the composited function
$\iota\circ\mu$ could map to only finite number of intervals. For
example, we can weaken the convex condition, Eq.~(\ref{eq:iota-Inclusion}),
to the following additive condition: for any numbers $x_{0}$ and
$x_{1}\in\left[0,1\right]$, their sum $x_{0}+x_{1}\in\left[0,1\right]$
implies
\begin{equation}
\iota\left(x_{0}+x_{1}\right)\subseteq\iota\left(x_{0}\right)+\iota\left(x_{1}\right)\,.\label{eq:iota-add}
\end{equation}
Then, given a three-dimensional Hilbert space, when $\iota$ satisfies
Eqs.~(\ref{eq:iota-constraints}) and Eq.~(\ref{eq:iota-add}),
$\iota\circ\mu$ still must be a QIVPM~\citep{HOSTunpublished}.
More concretely, consider two states $\rho_{0}=\frac{1}{3}\cdot\proj{0}+\frac{1}{3}\cdot\proj{1}+\frac{1}{3}\cdot\proj{2}$
and $\rho_{1}=\frac{7}{24}\cdot\proj{0}+\frac{1}{3}\cdot\proj{1}+\frac{3}{8}\cdot\proj{2}$,
and a set of intervals
\begin{equation}
\begin{aligned}\mathscr{I}_{2}={} & \left\{ \imposs,\left[0,\tfrac{1}{4}\right],\left[0,{\textstyle \tfrac{1}{2}}\right],\left[\tfrac{1}{4},\tfrac{3}{4}\right],\right.\\
 & \left.\left[\tfrac{1}{2},1\right],\left[\tfrac{1}{4},1\right],\necess\right\} \,.
\end{aligned}
\end{equation}
Since the interval-valued function~$\iota_{2}:\left[0,1\right]\rightarrow\mathscr{I}_{2}$
defined in Table~\ref{tab:I_2} satisfies Eqs.~(\ref{eq:iota-constraints})
and Eq.~(\ref{eq:iota-add})~\citep{HOSTunpublished}, $\iota_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$
and $\iota_{2}\circ\mu_{\rho_{1}}^{\mathrm{B}}$ are both QIVPMs.
\begin{table}
\caption{\label{tab:I_2}A interval-valued function~$\iota_{2}:\left[0,1\right]\rightarrow\mathscr{I}_{2}$.}
\centering{}%
\begin{tabular}{cc}
\toprule 
\addlinespace
$x$ & $\iota_{2}\left(x\right)$\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
$x=0$ & $\imposs$\tabularnewline\addlinespace
\midrule 
\addlinespace
$0<x\le\tfrac{1}{8}$ & $\left[0,\tfrac{1}{4}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\tfrac{1}{8}<x<\tfrac{3}{8}$ & $\left[0,{\textstyle \tfrac{1}{2}}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\tfrac{3}{8}\le x\le\tfrac{5}{8}$ & $\left[\tfrac{1}{4},\tfrac{3}{4}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\tfrac{5}{8}<x<\tfrac{7}{8}$ & $\left[\tfrac{1}{2},1\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\tfrac{7}{8}\le x<1$ & $\left[\tfrac{1}{4},1\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$x=1$ & $\necess$\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}
Because of $\rho_{0}\in\coreBorn\left(\iota_{2}\circ\mu_{\rho_{1}}^{\mathrm{B}},\events\right)$
and $\rho_{1}\in\coreBorn\left(\iota_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}},\events\right)$~\citep{HOSTunpublished},
Thm.~\ref{thm:Finite-precision-Gleason} asserts $\left\Vert \rho_{0}-\rho_{1}\right\Vert $
less than or equal to the maximum length of intervals $\frac{1}{2}$.
This is true because the exact value of $\left\Vert \rho_{0}-\rho_{1}\right\Vert $
is $\frac{1}{24}$.

>>>>

\begin{definition}[QIVPM]\label{def:QIVPM}
  Assume a collection of intervals $\mathscr{I}$ including $\imposs$
  and $\necess$ with addition and scalar multiplication defined as
  follows:
  \begin{subequations}\label{eq:interval-operations}
  \begin{eqnarray}
   &  & [\ell_{1},r_{1}]+[\ell_{2},r_{2}]=[\ell_{1}+\ell_{2},r_{1}+r_{2}]\textrm{ and}\\
   &  & x[\ell,r]=\begin{cases}
  [x\ell,xr] & \textrm{for }x\ge0\,;\\{}
  [xr,x\ell] & \textrm{for }x\le0\,.
  \end{cases}
  \end{eqnarray}
  \end{subequations}
  Then we take a QIVPM~$\bar{\mu}$ to be an assignment of an interval to each
  event (projection operator~$P$) subject to the following constraints:
  \begin{subequations}\label{eq:QIVPM-constraints}
  \begin{eqnarray}
   &  & \bar{\mu}(\mathbb{0})=\imposs\,,\\
   &  & \bar{\mu}(\mathbb{1})=\necess\,,\label{eq:necess}\\
   &  & \bar{\mu}\left(\mathbb{1}-P\right)= \necess-\bar{\mu}\left(P\right)\,,\label{eq:complement}
  \end{eqnarray}
  \end{subequations}
  and satisfying for each pair of \emph{commuting} projectors~$P_0$
  and~$P_1$ with $P_0P_1=P_1P_0$,
\begin{equation}
\bar{\mu}\left(P_{0}+P_{1}-P_{0}P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)\,.
\label{eq:QuantumInterval-valuedProbability-Inclusion}
\end{equation}
\end{definition}
\noindent The first three constraints,
Eqs.~(\ref{eq:QIVPM-constraints}), are the direct counterpart of the
corresponding ones for classical IVPMs.  With the understanding that
the union of classical sets $E_1\cup E_2$ is replaced by
$P_0+P_1-P_0P_1$ in the case of quantum projection
operators~\cite{Griffiths2003}, the last condition,
Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}), is a
direct counterpart of the convexity condition of
Eq.~(\ref{eq:classicalconvex}). Thus our definition of QIVPMs merges
classical IVPMs with quantum probability measures.

@@

Even if $\iota$ has to be a THOS rule to guarantee $\iota\circ\mu$
a QIVPM, the composited function $\iota\circ\mu$ may still be a QIVPM
when $\iota$ is not a THOS rule. In this case, the composited function
$\iota\circ\mu$ could map to only finite number of intervals. For
example, we can weaken the convex condition, Eq.~(\ref{eq:iota-Inclusion}),
to the following additive condition: for any numbers $x_{0}$ and
$x_{1}\in\left[0,1\right]$, their sum $x_{0}+x_{1}\in\left[0,1\right]$
implies
\begin{equation}
\iota\left(x_{0}+x_{1}\right)\subseteq\iota\left(x_{0}\right)+\iota\left(x_{1}\right)\,.\label{eq:iota-add}
\end{equation}
Then, given a three-dimensional Hilbert space, when $\iota$ satisfies
Eqs.~(\ref{eq:iota-constraints}) and Eq.~(\ref{eq:iota-add}),
$\iota\circ\mu$ still must be a QIVPM~\citep{HOSTunpublished}.
More concretely, consider two states $\rho_{0}=\frac{1}{3}\cdot\proj{0}+\frac{1}{3}\cdot\proj{1}+\frac{1}{3}\cdot\proj{2}$
and $\rho_{1}=\frac{7}{24}\cdot\proj{0}+\frac{1}{3}\cdot\proj{1}+\frac{3}{8}\cdot\proj{2}$,
and a set of intervals
\begin{equation}
\begin{aligned}\mathscr{I}_{2}={} & \left\{ \imposs,\left[0,\tfrac{1}{4}\right],\left[0,{\textstyle \tfrac{1}{2}}\right],\left[\tfrac{1}{4},\tfrac{3}{4}\right],\right.\\
 & \left.\left[\tfrac{1}{2},1\right],\left[\tfrac{1}{4},1\right],\necess\right\} \,.
\end{aligned}
\end{equation}
Since the interval-valued function~$\iota_{2}:\left[0,1\right]\rightarrow\mathscr{I}_{2}$
defined in Table~\ref{tab:I_2} satisfies Eqs.~(\ref{eq:iota-constraints})
and Eq.~(\ref{eq:iota-add})~\citep{HOSTunpublished}, $\iota_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$
and $\iota_{2}\circ\mu_{\rho_{1}}^{\mathrm{B}}$ are both QIVPMs.
\begin{table}
\caption{\label{tab:I_2}A interval-valued function~$\iota_{2}:\left[0,1\right]\rightarrow\mathscr{I}_{2}$.}
\centering{}%
\begin{tabular}{cc}
\toprule 
\addlinespace
$x$ & $\iota_{2}\left(x\right)$\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
$x=0$ & $\imposs$\tabularnewline\addlinespace
\midrule 
\addlinespace
$0<x\le\tfrac{1}{8}$ & $\left[0,\tfrac{1}{4}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\tfrac{1}{8}<x<\tfrac{3}{8}$ & $\left[0,{\textstyle \tfrac{1}{2}}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\tfrac{3}{8}\le x\le\tfrac{5}{8}$ & $\left[\tfrac{1}{4},\tfrac{3}{4}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\tfrac{5}{8}<x<\tfrac{7}{8}$ & $\left[\tfrac{1}{2},1\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\tfrac{7}{8}\le x<1$ & $\left[\tfrac{1}{4},1\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$x=1$ & $\necess$\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}
Given any one dimensional projector $\proj{\psi}$, we have $\mu_{\rho_{0}}^{\mathrm{B}}\left(\proj{\psi}\right)=\frac{1}{3}$
so that $\iota_{2}\left(\mu_{\rho_{0}}^{\mathrm{B}}\left(\proj{\psi}\right)\right)=\left[0,{\textstyle \tfrac{1}{2}}\right]$.
On the other hand, 
\begin{eqnarray*}
\mu_{\rho_{1}}^{\mathrm{B}}\left(\proj{\psi}\right) & = & \frac{7}{24}\cdot\left|\ip{\psi}{0}\right|^{2}+\frac{1}{3}\cdot\left|\ip{\psi}{0}\right|^{2}+\frac{3}{8}\cdot\left|\ip{\psi}{0}\right|^{2}
\end{eqnarray*}
which is bounded between $\frac{7}{24}$ and $\frac{3}{8}$. Therefore,
$\mu_{\rho_{1}}^{\mathrm{B}}\left(\proj{\psi}\right)\in\iota_{2}\left(\mu_{\rho_{0}}^{\mathrm{B}}\left(\proj{\psi}\right)\right)$.
According to the complement condition Eq.~(\ref{eq:complement}),
any two dimensional projector~$P$, we also have $\mu_{\rho_{1}}^{\mathrm{B}}\left(P\right)\in\iota_{2}\left(\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)\right)$.
Hence, $\rho_{1}\in\coreBorn\left(\iota_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}},\events\right)$,
and Thm.~\ref{thm:Finite-precision-Gleason} asserts $\left\Vert \rho_{0}-\rho_{1}\right\Vert $
less than or equal to the maximum length of intervals $\frac{1}{2}$.
This is true because the exact value of $\left\Vert \rho_{0}-\rho_{1}\right\Vert $
is $\frac{1}{24}$.
